Fast and optimal centroiding of faint stars

### 2015-02-16 ms/figure.tex
Update figure.tex

In this paper, we propose to find the centroid of stars by smoothing the image
of stars by a Gaussian kernel, and fitting a second order polynomial to
the $3\times$3 pixels around the brightest pixel of smoothed image.
We compare the accuracy of this method with, the naive 3$\times$3 second-order 
polynomial method, the 3$\times$3 quartic approximation
method used in SDSS photometric pipeline, and fitting a PSF model to the 
star.

In order to do so, we apply these methods to a large number of simulated
stars with different signal-to-noise ratio and size realizations. 
Error from star centroiding methods, will always have 
a theoretically-set lower bound, known as the Cramer-Rao lower bound
(hereafter denoted by CRLB) which has an inverse relation with the
signal-to-noise-ratio of stars.

Given the analytic model for the PSF,
we derive an expression for CRLB on the centroiding error as
a function of the parameters of the PSF model (\eg FWHM),
and signal-to-noise-ratio of stars. We create two sets of simulations for which we 
know the CRLB, one with variable (constant) signal-to-noise ratio (FWHM), and one 
with variable (constant) FWHM (signal-to-noise ratio). After applying
different centroiding methods to the simulations,
we investigate how close these methods can get to saturating the CRLB for
various ranges of background noise and FWHM.

@@

\subsection{3$\times$3 polynomial method}

In order to find the centroid of a star, 
we fit a simple 2d polynomial $P(x,y)=a+bx+cy+dx^2+exy+fy^2$ to the
$3\times3$ patch centered on the brightest pixel of the
image. We apply this method to star images in the sky-dominated regime.
In this regime, the covariance matrix of pixel uncertianties is simply
given by $\sigma^{2}\mathbf{I}$, where $\sigma^{2}$ is variance of the
uncorrelated Gaussian noise, and $\mathbf{I}$ is a 9$\times$9 identity matrix.
Upon constructing a universal 9$\times$6 design matrix
\begin{equation}
    \mathbf{A} = 
    \begin{bmatrix}
        1 & x_{1} & y_{1} & x_{1}^{2} & x_{1}y_{1} & y_{1}^{2} \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        1 & x_{9} & y_{9} & x_{9}^{2} & x_{9}y_{9} & y_{9}^{2}
    \end{bmatrix},
\end{equation}
the free parameters of the model $\{a,b,c,d,e,f\}$
(hereafter compactly denoted by $\mathbf{X}$) can be determined by 
\beq
\mathbf{X} = (\mathbf{A}^{T}\mathbf{A})^{-1}\mathbf{A}^{T}\mathbf{Z},
\label{linearfit}
\eeq
where $\mathbf{Z}$ is given by $(z_{1},...,z_{9})^{T}$,
with $z_{i}$, being the brightness of the $i-$th pixel of the 3$\times$3 patch.
Afterwards, the best fit parameters can be used to compute the centroid coordinate

@@

It is important to note that the algebraic operation in (\ref{center}) involves 
inverting a 2$\times$2 curvature matrix
\beq
  D = 
  \begin{bmatrix}
      2d & e\\
      e & 2f\\
  \end{bmatrix}.
\eeq
When D has a zero (or very close to zero) deteminant,
centroid estimated by (\ref{center}) gets arbitrarily 
large in which case the centroiding method leads to a catastrophic 
outlier. In order to tackle this issue, we add a small regularization term
proportional to $\sigma$ to the diagonal elements of the curvature matrix.
This soft regularization term prevents the determinant of the curvature matrix from
getting zero, and removes the catastrophic outliers.

\subsection{3$\times$3 modified polynomial method}
Now, we modify the simple polynomial method explained above,
by smoothing the image of the star with a Gaussian kernel
\beq
k(\mathbf{x}) = \frac{1}{2\pi w^2}\exp(-\mathbf{x}^{2}/2w^{2}).
\eeq
Throughout this study, we keep the width of the Gaussian kernel $w$ constant.
Then we apply the same polynomial method to the 3$\times$3 patch around the brightest
pixel of the smoothed image. The important difference between the previous
approach and the modified method is that, there are non-zero covariances
among the uncertainties of the brightness of different pixels of the
smoothed image. That is, the covariance matrix $\mathbf{C}$ is non-diagonal.
It can be shown that after smoothing, elements of the covariance matrix are given by
\begin{eqnarray}
\mathbf{C}_{ij} &=& \frac{\sigma^{2}}{4\pi w^{2}} \exp(-\frac{r_{ij}^{2}}{4\pi w^{2}}),\\
r_{ij}^{2} &=& (x_{i} - x_{j})^{2} + (y_{i} - y_{j})^{2},
\label{nondiagonal}
\end{eqnarray}
where $w$ is width of the kernel, $\sigma$ is the per-pixel uncertainty, and $x_{i}$
, $y_{i}$ are the $x$, and $y$ coordinates of the $i$-th pixel
respectively. In the presence of the non-diagonal covariance matrix
$\mathbf{C}$, equation (\ref{linearfit}) needs to be modified in the following way
\beq
\mathbf{X} = [\mathbf{A}^{T}\mathbf{C}^{-1}\mathbf{A}]^{-1}[\mathbf{A}^{T}\mathbf{C}^{-1}\mathbf{Z}].
\label{linearfit2}
\eeq
Therefore, for a given star and a smoothing kernel,
the outcome of equation (\ref{linearfit2}) can be
plugged into equation (\ref{center}) to find the centroid estimate
of the star.
\subsection{3$\times$3 quartic approximation method}
Consider any row or column of the 3$\times$3 patch around
the brightest pixel of the smooth image and let us denote
the brightness of these three pixels by $\{I_{-},I_{0},I_{+}\}$.
The fundamental assumption behind the technique is that these
brightness values can be approximated
by the function $P(x) = Q\exp((x-x_{c})^{2}/2\beta^{2})$
evaluated at $x=-1,0,+1$; with $x_{c}$ being the position of centroid
of the three pixels. Knowing the value of $P$ at $\{-1,0,1\}$,
expanding $P(x)$ up to fourth order in $x$ leads to
\beq
x = \frac{s}{t}(1+k\frac{t}{4Q}(1-(s/t)^{2})),
\eeq
where k = 1.33, and

\begin{eqnarray}
s&=&(I_{+}-I_{-})/2, \\
t &=&2I_{0} - (I_{+}+I_{-}), \\
Q &=& I_{0} +s^{2}/2t.
\label{def}
\end{eqnarray}

If we apply this procedure to every column and row of the 3$\times$3 patch
around the brightest pixel of the smoothed image, we will find three local
maxima for three rows $\{(x_{-},-1)$, $(x_{0},0)$, $(x_{+},1)\}$, and three
local maxima for three columns $\{(-1,y_{-})$, $(0,y_{0})$, $(1,y_{+})\}$.
The centroid is then estimated by finding the intersection of two curves,
one fitted to three maxima corresponding to rows, and the other fitted
to three maxima corresponding to columns.

\subsection{7$\times$7 weighted moments}

In addition to the methods mentioned above, we examine weighted moment method which
requires taking a 7$\times$7 patch of image around the brightest pixel,
weighting the pixels by the intensity, and computing the first moments (or center
of mass) of the weighted 7$\times$7 patch thereafter.

\section{Simulations and Tests}\label{sec:data}

In order to test the accuracy of the method mentioned so far, and to ompare them against each other,
and we need to simulate a large set of stars for which we know the exact position of centroids.
Furthermore, we uniformly draw the centroids within sub-pixel regions. 
We use the Moffat profile \citep{moffat} for our PSF simulations. 
Moffat profile is an analytic model for stellar PSFs. It has broader wings than
a simple Gaussian model. The surface brightness of the Moffat profile is given by
\beq
I(r) = \frac{F(\beta -1)}{\pi \alpha^{2}}[1+(r/\alpha)^{2}]^{-\beta},
\label{mof}
\eeq
where $F$ is the total flux, $\beta$ is a dimensionless paramter, and $\alpha$ is
the length scale of the Moffat profile, with FWHM (hereafter denoted by $\gamma$)
being $2\alpha\sqrt{2^{1/\beta}-1}$. 
We want to investigate the performance of centroiding methods introduced
in section \ref{sec:method} for different background noise levels, and also different
values of $\gamma$. For further simplicity, we set the flux of all stars in our
simulations to unity. Per pixel uncertainties are assumed to be uncorrelated Gaussian,
and we only study the sky-dominated regime.

Moreover, it is more convenient to work with the signal-to-noise ratio
(hereafter denoted by SNR) instead of the variance of the Gaussian noise.
We define SNR as the ratio of the mean and variance of the distribution
which the flux estimator is drawn from. Assume that the total flux from
the point source is $f$, and that the PSF at the $i$-th pixel is given
by $P_{i}$. Therefore the brightness of the $i$-th pixel is drawn from
a Gaussian distribution $p(I_{i}) = \mathcal{N}(fP_{i},\sigma^{2})$.

The optimal estimator of flux (one which saturates Cramer-Rao bound),
is the weighted average $\tilde{f}=\sum_{i}I_{i}P_{i}$. Knowing the
distribution of $I_{i}$, it can be shown that 
\beq
p(\tilde{f}) = \mathcal{N}(f , \frac{\sigma^{2}}{\sum_{i}P_{i}^{2}}),
\eeq  
which leads us to
\beq
\begin{array}{l}
\text{SNR} $=$ \frac{\sqrt{\sum_{i} P_{i}^{2}}}{\sigma}.
\end{array}
\label{snr}
\eeq

Note that (\ref{snr}) is valid in the limit of sky-dominated regimes,
and that is the limit we are interested in. In the case of our simulations,
SNR given in (\ref{snr}) can be analytically expressed in terms of the per pixel uncertainty
$\sigma$, FWHM $\gamma$, and also $\beta$. We fix $\beta$ at the fiducial value of 2.5.
In this case we have
\beq
\text{SNR} = \frac{0.478}{\sigma \gamma}.
\label{snr2}
\eeq

Equation (\ref{snr2}) implies that for stars with the same flux $f$ and background
noise level $\sigma$, those with radius have lower SNR.

We aim to study how centroiding errors from different methods behave with
changing SNR and size, and how close they get to saturating the CRLB. 
Given the PSF model that we have adopted for our simulations, We need 
to derive an expression for the CRLB as a function of the size of of
the stars, and SNR. We know that the CRLB is given by the inverse of 
the Fisher information matrix $\mathcal{F}$. The closer an estimator is
to saturating the CRLB, the more information about the quantity that we 
need to estimate is preserved. A measure for closeness of an estimator to 
saturating the CRLB is comparison of the root-mean-squared-error of that estimator 
with the CRLB.

In order to find the CRLB, it is sufficient to find the Fisher matrix.
Let us assume that we have $M$ observables $\{f_{1}, ... , f_{M}\}$, each
related to $B$ model parameters $\{\theta_{1} , ... , \theta_{B}\}$. Assuming
uncorrelated Gaussian error $\sigma_{m}$ for each observable $f_{m}$, elements
of the Fisher matrix $\mathcal{F}_{ij}$ are given by
\beq
\mathcal{F}_{ij} = \sum_{m}\frac{1}{\sigma_{m}^{2}}\frac{\partial f_{m}}{\partial \theta_{i}}\frac{\partial f_{m}}{\partial \theta_{j}}
\label{fisher}
\eeq

@@

Equation (\ref{crlbmoffat}) implies that at given SNR and $\gamma$
CRLB for each component of centroid is given by $0.6846\gamma/\text{SNR}$,
and that a good centroiding technique delivers centroids with
root-mean-squared-error (hereafter RMSE) of each component close to
this number.

We perform two sets of simulations. In the first set, we choose four values of
2, 2.8, 4, and 5.6 pixels for $\gamma$. For each $\gamma$, we generate 100,000 
17 $\times$ 17 postage-stamps of Moffat profiles with centroids randomly drawn
within the central pixel of the 17 $\times$ 17 postage-stamps. Moreover, Gaussian
is noise added to each postage-stamp such that the stars are uniformly distributed
in log-SNR between SNR = 5 to SNR = 100.

In the second set, we generate 100,000 17$\times$17 postage-stamps
of Moffat profile, with values of $\gamma$ randomly distributed 
between 2 pixels and 6 pixels, and with centroids drawn randomly within 
the central pixel. we choose four values for SNR: 5, 10, 20, and 40. 
For each SNR, and for each postage-stamp with a given $\gamma$, 
Gaussian noise, with variance corresponding to SNR and $\gamma$ through equation 
(\ref{snr2}), is added to the each postage-stamp.

In the first experiment, we study how centroiding error behaves with changing
SNR, while $\gamma$ is being held constant. In the second experiment, we study 
how centroiding error behaves with changing $\gamma$ while SNR is being held constant.

In addition to the centroiding methods discussed in the previous sections,
we compute the error from fitting the Moffat PSF model with the exact $\gamma$ to
the data. That is, we find the best estimates of flux and centroid by
optimizing the $\chi^{2}$. Furthermore, we computeth error from fitting a
wrong PSF model (Gaussian profile) with the exact $\gamma$ to the data.
In all cases in both experiments, we expect the PSF fitting method to saturate 
the CRLB.

@@

In this experiment, after finding the centroiding error for each method,
we compute the RMSE in bins of SNR in order to compare it to the CRLB. 
Results of the first experiment are shown in Figures [\ref{1}], [\ref{2}],
[\ref{3}], [\ref{4}], [\ref{5}], [\ref{6}]. All methods deliver results
with RMSE larger for fainter stars.
In case of 3$\times$3 polynomial centroiding [\ref{1}], RMSE becomes
very large as we move towards the fainter stars in our simulation.
Moreover, we observe a kink in the plot of RMSE against SNR.
For stars with larger $\gamma$, this kink moves toward higher values of SNR.
As it can be seen in figure [\ref{1}], centroid estimate from the polynomial
method do not even get close to saturating the CRLB. By increasing $\gamma$,
RMSE moves further away from the CRLB.

However, RMSE from the modified centroiding method [\ref{2}],
and RMSE from the quartic approximation [\ref{3}] method are very
close to the CRLB except at the very small values of SNR (SNR $<$ 10).
As we increase $\gamma$ from 2 pixels to 2.8 pixels, RMSE gets closer
to the CRLB. This implies that as stars get undersampled, the performance
of these two models get slightly worse. But if we increase $\gamma$ to
4 and 5.6 pixels, RMSE gets farther from the CRLB.
This effect is quite predictable since these two models only use the
information contained in a 3$\times$ 3 patch. It is also important to
note that the rate at which the RMSE from these two method drops
eventually becomes smaller than the constant rate at which the CRLB
decreases with increaing SNR. Also the results from 7$\times$7 weighted moment
method [\ref{4}] suggest that this method does worse than the qurtic approximation, and
the modified polynomial method, and at low values of SNR the centroiding RMSE from
this method moves significantly away from the CRLB.

As we expected, the RMSE from centroiding by fitting the exact PSF model [\ref{5}]
lies on the CRLB except at values of SNR below 10 where the RMSE gets slightly pulled away
from the CRLB because of the presence of a few outliers.
Thus, centroid estimates from fitting the exact PSF model always
saturate the CRLB. As it can be seen in figure [\ref{6}],
even the centroid estimates found by fitting a wrong PSF model (Gaussian PSF)
to the simulated stars saturate the CRLB.

\subsection{Experiment 2}
In this experiment, after finding the centroiding error for each method, we
compute the RMSE in bins of $\gamma$ in order to compare it to the CRLB. 
Behavior of error as a function of $\gamma$ for different values of SNR,
is shown in figures [\ref{7}], [\ref{8}], and [\ref{9}], [\ref{10}], [\ref{11}], and [\ref{12}].

The centroid estimates obtained from the 3$\times$3 polynomial method [\ref{7}]
result in RMSE much larger than the CRLB in all ranges of FWHM and for all four
values of SNR in this experiment. In the case of extremely faint stars (SNR = 5),
the polynomial method is not able to deliver any reliable estimate, and completely fails.

The 3$\times$3 modified polynomial method [\ref{8}], and the 3$\times$3
quartic approximation [\ref{9}] result in RMSE very close to the CRLB.
For al four values of SNR, as we increase $\gamma$ from 2 pixels to 3 pixels,
RMSE gets slightly closer to the CRLB since these two models start to work
slightly worse as we decrease $\gamma$ and move towards undersampled stars.
After approximately 3 pixels, increasing $\gamma$ deviate RMSE of these methods from the CRLB.
This is inherent characteristic of all 3$\times$3 methos. 
Furthermore, increasing SNR from 5 to 40 makes the RMSE (as a function of $\gamma$) get
closer to the CRLB. Also we observe that in the case of 7$\times$7 weighted moment 
centroiding, deviation of the centroiding RMSE from the CRLB is significantly higher
than those of 3$\times$3 quartic approximation, and 3$\times$3 modified polynomial.

Again The RMSE from centroiding by fitting the exact PSF model as a function of FWHM
perfectly lies on the CRLB [\ref{11}] except at SNR = 5 where the RMSE slightly deviates
from the CRLB due tp the presence of outliers. Thus, centroid estimates from fitting
the exact PSF model always saturate the CRLB. Once again, we observe that the centroid
estimates found by fitting a Gaussian PSF to the simulated stars also saturate
-- except in the case of SNR = 5-- the CRLB in all ranges of $\gamma$ [\ref{12}].

@@

An optimal stellar centroiding algorithm must always saturate the Cramer-Rao lower 
bound. That is, in all ranges of background noise level, size, radial light profile,
and shape, it must preserve information about the centroid of stars. In practice however,
that is only achievable when we know the exact PSF. Since, we do not know the exact PSF
profile, we make use of various approximate centroiding algorithms. In this work, we
studied how different methods get close to saturating the CRLB for relatively low SNR
stars.

We also proposed a method for centroiding stars based on fitting a low order
polynomial to a 3$\times$3 patch of star images smoothed by a Gaussian kernel. 
Furthermore, we compared the performance of this method with the naive 3$\times$3 polynomial
centroiding, 3$\times$3 quartic approximation, moment-based method, and PSF fitting.
In order to make this comparison, we performed some tests on a large number of
simulated data, and investigated to see if these estimators can saturate the Cramer-Rao
lower bound, or how close to they get to saturating the bound.

Our results suggest that in all ranges of FWHM and SNR, the PSF fitting method returns 
centroid estimates that saturate the CRLB; with the exception of SNR being less than 10, 
in which a few centroiding ouliers caused the RMSE to slightly, but not significantly, deviate
from the CRLB. We found that the estimates found by 3$\times$3 polynomial method are far away from
saturating the CRLB. On the other hand, the RMSE of centroid estimates of our proposed
3$\times$3 modified polynomial method, and that of the 3$\times$3 quartic approximation,
are much closer to the CRLB. The performance of these two methods however, is limited by
two important factors: \emph{(i)} signal-to-noise ratio, and \emph{(ii)} size of the stars.

>>>>

In this paper, we study how optimal various techniques are at centroiding 
faint stars. That is, we apply a number of centroiding methods 
to a large number of simulated stars with different signal-to-noise ratio and
 size realizations. Error from star centroiding methods, will always have 
a theoretically-set lower bound, known as the Cramer-Rao lower bound
(hereafter denoted by CRLB) which has an inverse relation with the
signal-to-noise-ratio of stars.

Given the analytic model for the PSF,
we derive an expression for the fundamental lower bound on the centroiding error as
a function of the parameters of the PSF model (\eg size),
and signal-to-noise-ratio of stars. We create two sets of simulations for which we 
know the CRLB, one with variable (constant) signal-to-noise ratio (FWHM), and one 
with variable (constant) FWHM (signal-to-noise ratio). After applying
different centroiding methods to the simulations,
we investigate how close these methods can get to saturating the CRLB for
various ranges of background noise and FWHM.

In this work, we focus on three centroiding methods. The first method finds
 the centroid of stars by smoothing the image of
 stars by a Gaussian kernel of a fixed size, and fitting a second order
 polynomial to the $3\times$3 patch around the brightest pixel of smoothed
 image. This method is fast and does not require any knowledge of the 
PSF. In the second method, we apply the same polynomial centroiding to the
 $3\times$3 patch around the brightest pixel of the image of star convolved 
with the full PSF postage stamp. The last method that we consider, is fitting 
a PSF profile, assuming that we have a good PSF estimate, to the star.
%We compare the accuracy of this method with, the naive 3$\times$3 second-order 
%polynomial method, the 3$\times$3 quartic approximation
%method used in SDSS photometric pipeline, and fitting a PSF model to the 
%star.

@@

\subsection{Non PSF-based methods}

\item {\bf 3$\times$3 polynomial centroiding}\quad

In order to find the centroid of a star, 
we fit a simple 2d polynomial $P(x,y)=a+bx+cy+dx^2+exy+fy^2$ to the
$3\times3$ patch centered on the brightest pixel of the
image. We apply this method to star images in the sky-dominated regime.
Upon constructing a universal 9$\times$6 design matrix
\begin{equation}
    \mathbf{A} = 
    \begin{bmatrix}
        1 & x_{1} & y_{1} & x_{1}^{2} & x_{1}y_{1} & y_{1}^{2} \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        1 & x_{9} & y_{9} & x_{9}^{2} & x_{9}y_{9} & y_{9}^{2}
    \end{bmatrix},
\end{equation}
the free parameters of the model $\{a,b,c,d,e,f\}$
(hereafter compactly denoted by $\mathbf{X}$) can be determined by 
\beq
\mathbf{X} = (\mathbf{A}^{T}\mathbf{A})^{-1}\mathbf{A}^{T}\mathbf{Z},
\label{linearfit}
\eeq
where $\mathbf{Z}$ is given by $(z_{1},...,z_{9})^{T}$,
with $z_{i}$, being the brightness of the $i-$th pixel of the 3$\times$3 patch.
Afterwards, the best fit parameters can be used to compute the centroid coordinate

@@

It is important to note that the algebraic operation in (\ref{center}) involves 
inverting a 2$\times$2 curvature matrix
\beq
  D = 
  \begin{bmatrix}
      2d & e\\
      e & 2f\\
  \end{bmatrix}.
\eeq
When $D$ has a zero (or very close to zero) deteminant,
centroid estimated by (\ref{center}) gets arbitrarily 
large in which case the centroiding method leads to a catastrophic 
outlier. In order to tackle this issue, we add a soft regularization term
proportional to $\sigma^{2}$ to the determinant of $D$.

\item {\bf 3$\times$3 modified polynomial method I}\quad

Now, we modify the simple polynomial method explained above,
by smoothing the image of the star with a Gaussian kernel
\beq
k(\mathbf{x}) = \frac{1}{2\pi w^2}\exp(-\mathbf{x}^{2}/2w^{2}).
\eeq
Throughout this study, we keep the width of the Gaussian kernel $w$ constant and equal to
1.2 pixels (corresponding to FWHM = 2.8 pixels).
Then we apply the same polynomial method to the 3$\times$3 patch around the brightest
pixel of the smoothed image. Therefore, for a given star and a smoothing kernel,
the outcome of equation (\ref{linearfit}) can be
plugged into equation (\ref{center}) to find the centroid estimate
of the star.

\subsection{PSF-based methods}

\item {\bf 3$\times$3 modified polynomial method II}\quad

Consider the case that we have a good estimate of the PSF at
the position of the faint star under consideration. Therefore, we 
can modify the smoothing step in the following way. Instead of smoothing
with a Gaussian kernel with a fixed size, we convolve the image of star
with the full postage stamp of the PSF. Then, we apply the same polynomial
fitting to the 3$\times$3 patch around the brightest pixel of the convolved image.

The procedure of convolving the image of star with the PSF results in a
more well-sampled, which eventually, leads to having more
information regarding the centroid of stars in the 3$\times$3 patch around the
brightest pixel. In addition, a simple 2d polynomial will provides a better fit 
since convolution with the PSF makes the variation of the image across the 3 pixels
very smooth.

\item {\bf psf profile fitting}\quad

In addition to the methods mentioned so far, we examine fitting stars with the exact
 PSF profile. That is, in our Cramer-Rao bound saturation tests, That is, we find the best
 estimates of flux and centroid by
optimizing the $\chi^{2}$. We expect this method to perform better in determing the centroid
 of stars. In theory, it always saturates the CRLB.

\section{Simulations and Cramer-Rao saturation tests}\label{sec:data}

In order to test the accuracy of the method mentioned in section \ref{sec:method},
 and to compare their performance at saturating the Cramer-Rao bound against each other, 
we need to simulate a large set of stars for which we know the exact position of centroids.
Furthermore, we uniformly draw the centroids of these simulated stars within sub-pixel regions. 
We use the Moffat profile \citep{moffat} for our PSF simulations. 
Moffat profile is an analytic model for stellar PSFs. It has broader wings than
a simple Gaussian profile. The surface brightness of the Moffat profile is given by
\beq
I(r) = \frac{F(\beta -1)}{\pi \alpha^{2}}[1+(r/\alpha)^{2}]^{-\beta},
\label{mof}
\eeq
where $F$ is the total flux, $\beta$ is a dimensionless paramter, and $\alpha$ is
the length scale of the Moffat profile, with FWHM (hereafter denoted by $\gamma$)
being $2\alpha\sqrt{2^{1/\beta}-1}$. 
We aim to investigate the performance of centroiding methods introduced
in section \ref{sec:method} for different background noise levels, and also different
values of $\gamma$. For further simplicity, we set the flux of all stars in our
simulations to unity. Per pixel uncertainties are assumed to be uncorrelated Gaussian,
and we only study the sky-dominated regime.

Moreover, it is more convenient to work with the signal-to-noise ratio
(hereafter denoted by SNR) instead of the variance of the Gaussian noise.
We define SNR as the ratio of the mean and variance of the distribution
which the flux estimator is drawn from. Assume that the total flux from
the point source is $F$, and that the PSF at the $i$-th pixel is given
by $P_{i}$. Therefore the brightness of the $i$-th pixel is drawn from
a Gaussian distribution $p(I_{i}) = \mathcal{N}(FP_{i},\sigma^{2})$.

The optimal estimator of flux (one which saturates Cramer-Rao bound),
is the weighted average $\tilde{F}=\sum_{i}I_{i}P_{i}$. Knowing the
distribution of $I_{i}$, it can be shown that 
\beq
p(\tilde{F}) = \mathcal{N}(F , \frac{\sigma^{2}}{\sum_{i}P_{i}^{2}}),
\eeq  
which leads us to
\beq
\begin{array}{l}
\text{SNR} $=$ \frac{\sqrt{\sum_{i} P_{i}^{2}}}{\sigma}.
\end{array}
\label{snr}
\eeq

Note that (\ref{snr}) is valid in the limit of sky-dominated regimes,
and that is the limit we are interested in throughout our investigation.
 In the case of our simulations, SNR given in (\ref{snr}) can be analytically
 expressed in terms of the per pixel uncertainty
$\sigma$, FWHM $\gamma$, and also $\beta$. We fix $\beta$ at the fiducial value of 2.5.
In this case we have
\beq
\text{SNR} = \frac{0.478}{\sigma \gamma}.
\label{snr2}
\eeq

Equation (\ref{snr2}) implies that for stars with the same flux $F$,
 and background
noise level $\sigma$, those with higher size have lower SNR.

We aim to study how centroiding errors from different methods behave with
changing SNR and size, and how close they get to saturating the CRLB. 
The closer an estimator is to saturating the CRLB, the more information
 about the quantity that we 
need to estimate is preserved. A measure for closeness of an estimator to 
saturating the CRLB is comparison of the root-mean-squared-error of that estimator 
with the CRLB.

Given the PSF model that we have adopted for our simulations, we need 
to derive an expression for the CRLB as a function of the size, and SNR of stars.
 We know that the CRLB is given by the inverse of 
the Fisher information matrix $\mathcal{F}$. Thus, in order to find the CRLB,
 it is sufficient to compute the Fisher matrix.
Let us assume that we have $M$ observables $\{f_{1}, ... , f_{M}\}$, each
related to $B$ model parameters $\{\theta_{1} , ... , \theta_{B}\}$. Assuming
uncorrelated Gaussian error $\sigma_{m}$ for each observable $f_{m}$, elements
of the Fisher matrix $\mathcal{F}_{ij}$ are given by
\beq
\mathcal{F}_{ij} = \sum_{m}\frac{1}{\sigma_{m}^{2}}\frac{\partial f_{m}}{\partial \theta_{i}}\frac{\partial f_{m}}{\partial \theta_{j}}
\label{fisher}
\eeq

@@

Equation (\ref{crlbmoffat}) implies that at given SNR and $\gamma$,
CRLB for each component of centroid is given by $0.6846\gamma/\text{SNR}$,
and that a good centroiding technique delivers centroids with
root-mean-squared-error (hereafter RMSE) of each component close to
this number.

We perform two sets of simulations. In the first set, we choose four values of
2, 2.8, 4, and 5.6 pixels for $\gamma$. For each $\gamma$, we generate 100,000 
17 $\times$ 17 postage-stamps of Moffat profiles with centroids randomly drawn
within the central pixel of the 17 $\times$ 17 postage-stamps. Moreover, Gaussian
noise is added to each postage-stamp such that the simulated stars are uniformly
 distributed in log-SNR between SNR = 5 to SNR = 100.

In the second set, we generate 100,000 17$\times$17 postage-stamps
of Moffat profile, with values of $\gamma$ randomly distributed 
between 2 pixels and 6 pixels, and with centroids drawn randomly within 
the central pixel. We choose four values for SNR: 5, 10, 20, and 40. 
For each SNR, and for each postage-stamp with a given $\gamma$, 
Gaussian noise, with variance corresponding to SNR and $\gamma$ through equation 
(\ref{snr2}), is added to the each postage-stamp.

In the first experiment, we study how centroiding error behaves with changing
SNR, while $\gamma$ is held constant. In the second experiment, we study 
how centroiding error behaves with changing $\gamma$ while SNR is held constant.

@@

In this experiment, after finding the centroiding error for each method,
we compute the RMSE in bins of SNR in order to compare it to the CRLB. 
Results of the first experiment are shown in Figures [\ref{1}], [\ref{2}],
[\ref{3}], [\ref{4}]. All methods deliver results
with RMSE larger for fainter stars.
In case of 3$\times$3 polynomial centroiding [\ref{1}], RMSE becomes
quite large as we move towards fainter stars in our simulation.
For stars with larger $\gamma$, centroid estimate from the naive polynomial
method show large deviation from the CRLB. By increasing $\gamma$,
RMSE deviates further away from the CRLB.

However, RMSE from the modified centroiding method I [\ref{2}],
is very close to the CRLB except at the very small values of SNR (SNR $<$ 10).
As we increase $\gamma$ from 2 pixels to 2.8 pixels, RMSE gets closer
to the CRLB. Besides, Also, when we increase $\gamma$ to
4 and 5.6 pixels, RMSE gets farther from the CRLB. For stars with $\gamma$ = 2 pixels, 
the rate at which the RMSE from this method drops
eventually becomes smaller than the constant rate at which the CRLB
decreases with increaing SNR. The reason for this is that even after smoothing
the with a Gaussian kernel, the image is still not well-sampled, and smooth enough
for a 2d polynomial fitting to provide highly accurate centroiding estimates.
For stars with $\gamma=2.8$ pixels, since the FWHM of the Gaussian kernel matches that 
of the underlying simulations, the smooth images are well-sampled and therefore, 
the method is able to deliver estimates very close to saturating the CRLB. As $\gamma$
gets higher for the simulations, the convolved images are not as well-sampled as those 
whose $\gamma$ matches the FWHM of the smoothing kernel and as a result, we loose
some information by fitting a polynomial to only a 3$\times$3 patch. That is the 
reason for why we observe slight deviation from the CRLB for stas with $\gamma$ = 4 pixels, and 
slightly more deviations as we increase $\gamma$ to 5.6 pixels.

As we expected, the RMSE from centroiding by fitting the exact PSF model [\ref{4}]
lies on the CRLB except for simulations with SNR = 5 where the RMSE gets slightly pulled away
from the CRLB due to presence of a few outliers. Figure [\ref{3}] demonstrates that even the modified centroiding
 method II is able deliver centroiding estimates as accurate as PSF profile fitting, with the exception
of simulated stars with $\gamma$ = 2 pixels. For stars with $\gamma$ = 2 pixels, the 3$\times$3 polynomial
fitting fails to saturate the CRLB since the PSF is undersampled. For simulated images higher $\gamma$, convolving the
data with the PSF result in images that are well-sampled around the brightest pixel. This allows the 
polynomial centroiding to deliver highly accurate results that can saturate the CRLB for simulated
data with SNR $\ga$ 10.

\subsection{Experiment 2}

In this experiment, after finding the centroiding error for each method, we
compute the RMSE in bins of $\gamma$ in order to compare it to the CRLB. 
Behavior of error as a function of $\gamma$ for different values of SNR,
is shown in figures [\ref{5}], [\ref{6}], and [\ref{7}], and [\ref{8}].

The centroid estimates obtained from the naive 3$\times$3 polynomial method [\ref{5}]
result in RMSE much larger than the CRLB in all ranges of FWHM and for all four
values of SNR in this experiment. In the case of extremely faint stars (SNR = 5),
the polynomial method is not able to deliver any reliable estimate, and completely fails.

The 3$\times$3 modified polynomial method I [\ref{6}], results in RMSE very close to the CRLB.
For all four values of SNR, as we increase $\gamma$ from 2 pixels to 3 pixels,
RMSE gets slightly closer to the CRLB since the method starts to perform
slightly better as we move away from undersampled stars, and as the FWHM of the smoothing kernel
gets closer to that of simulated images.
After approximately 3 pixels, increasing $\gamma$ result in deviation of RMSE of the method from the CRLB.
This is a characteristic of 3$\times$3 method as we apply it to a smoothed image which is still not sufficiently
well sampled.  Furthermore, increasing SNR from 5 to 40 makes the RMSE (as a function of $\gamma$) get
closer to the CRLB.

Again The RMSE from centroiding by fitting the exact PSF model as a function of FWHM
perfectly lies on the CRLB [\ref{8}] except at SNR = 5 where the RMSE slightly deviates
from the CRLB due to presence of a few outliers. Thus, centroid estimates from fitting
the exact PSF model always saturate the CRLB. Once again, we observe that the centroid
estimates found by the 3$\times$3 modified polynomial method II saturate the CRLB with the
exception of simulated stars with SNR = 5, or $\gamma$ very close to 2 pixels. [\ref{7}].

@@

A perfectly optimal stellar centroiding algorithm must saturate the fundamental Cramer-Rao lower 
bound. That is, in all ranges of background noise level, size, radial light profile,
and shape, it must preserve information about the centroid of stars. In practice however,
that is only achievable when we have a very good estimate of the PSF. Since, we do not always 
know the exact PSF profile, we make use of various approximate centroiding algorithms. In this work, we
studied how different methods get close to saturating the CRLB for relatively low signal-to-noise ratio
stars.

We focused on examples from two classes of centroiding algorithms. The first class contains fast and approximate
methods that do not require any knowledge of the PSF at the poistion of stars. Of methods that belong to this class
we consider a simple method for centroiding stars based on fitting a low order
polynomial to a 3$\times$3 patch of star images, once smoothed by a Gaussian kernel of fixed width, and once without any smoothing step. 
The second class of centroiding algorithms contains methods which require knowledge of the PSF (or having a good estimate
of the PSF) at the position of stars. We considered two examples from this class. The first example
was the 3\times3 polynomial centroiding applied to the image of the star convolved with the PSF, and the
second example is the PSF profile fitting. In terms of saturating the Cramer-Rao bound we compared 
the performances of these mthods against each other. We call the 3$\times$3 polynomial method applied 
to the image convolved with a fixed Gaussian kernel (the PSF), 3$\times$3 polynomial method I (II).
